ICT and Future Productivity \\ Evidence and Theory of a GPT

### 2018-11-13 Latex/MacroLunch_Nov2018/Slides_Nov2018.tex
edited presentation based on yesterday's run-thru
didn’t change variables though, to stay consistent with IR-matching

\title[ICT \& medium-run TFP]{ICT and Future Productivity: Evidence and Theory of a GPT}

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Motivation}

@@

	\item The Great Recession 2008 casts doubt on whether this is all of the story (Fernald et al. (2017)).

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Wrong timing}

		\item [$\rightarrow$] A suspect is investment in information and communication technologies (ICT).

@@

	\begin{enumerate}
	\item Stick ICT investment (ICT-I) in a VAR and explore how an identified shock to this affects TFP.
	\item [] ICT shock = a shock to the productivity of the ICT sector today

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Baseline identification}
	\label{baseline_spec}

\begin{equation}\label{eq:mainObjective}
\max_{\gamma_j} \Pi_{2,j}(0) = e_2' \tilde{A}_0 \gamma_j
\end{equation}
subject to
\begin{equation}\label{eq:mainZeroTFP}
\Pi_{1,j}(0) = e_1' \tilde{A}_0 \gamma_j = 0, \ \ \text{and}
\end{equation}
\begin{equation}\label{eq:mainOrtho}
\gamma_j' \gamma_j = 1
\end{equation}
where $j$ represents the arbitrary position of the ICT shock.

	\begin{itemize}
	 \item[(2)] maximal impact effect on ICTI. 
	 \item[(3)] no impact effect on TFP. 
	 \item[(4)]$\gamma_j$ comes from an orthogonal matrix.

\hyperlink{VAR_notation}{\beamerbutton{VAR}}

@@

\end{frame}
%%%%%%%%%%%%%%%%%
%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Our favorite specification}

	\begin{itemize}
	\item Recall: dataset is quarterly and covers 1989:q1-2017-q2.

	\item One lag (as suggested by BIC and HQ).

	\item Horizon of FEV-maximization: 60 quarters.

	%\item Restriction on relative prices after a news shock is imposed at 8 quarters.

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Uses of outputs and GDP}

@@

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{TFP in the model}

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Impulse-response matching - does data support $\gamma>0$?}

@@

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{The spillover drives the TFP-response}
\begin{figure}[h!]
\begin{center}
\includegraphics[scale=0.13]{\MainFigures/fig_ICT_Shock_on_TFP_IRmatching_together_gam_set_to0point1}
\caption{Setting $\gamma =0.1$}
\label{fig:TFP_main}
\end{center}
\end{figure}

@@

@@

>>>>

\title[ICT \& medium-run TFP]{ICT and Future Productivity \\
Evidence and Theory of a GPT}

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Motivation}
	\label{motivation}

@@

	\item Fernald et al. (2017): the Great Recession 2008 casts doubt on whether this is all of the story.

@@

\hyperlink{wrong_timing}{\beamerbutton{Argument}}

@@

	\begin{enumerate}
	\item Stick aggregate ICT investment (ICT-I) in a VAR and explore how an identified shock to this affects TFP.
	\item [] ICT shock = a shock to the productivity of the ICT sector today

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Baseline identification: Cholesky}
	\label{baseline_spec}

	The ICT shock is assumed to have

	\begin{itemize}
	\item no impact effect on TFP;

	\item maximal impact effect on ICT-I.
	\end{itemize}

%\hyperlink{VAR_notation}{\beamerbutton{VAR}}

@@

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Uses of outputs}
	\label{uses}

@@

\hyperlink{tfp_in_model}{\beamerbutton{TFP in the model}}

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Impulse-response matching - does data support $\gamma>0$?}
	\label{IRmatching}

@@

and calibrate the rest.  \hyperlink{calibration}{\beamerbutton{Calibration}}

@@

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Wrong timing}
	\label{wrong_timing}

\hyperlink{motivation}{\beamerreturnbutton{Return}}

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{TFP in the model}
	\label{tfp_in_model}

\hyperlink{uses}{\beamerreturnbutton{Return}}

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{GDP and relative prices in the model}
	\label{tfp_in_model}

\hyperlink{uses}{\beamerreturnbutton{Return}}

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Calibration}
	\label{calibration}

\begin{table}[h!]
\footnotesize
	\begin{center}
		\begin{tabular}{clcl}
			\hline
			 & Economic Interpretation &  Value & Reference\\
			\hline
            $\beta$ & Discount factor of households &   $0.97^{\frac{1}{4}}$ & Match 3M-T-bill return \\
            $\delta^c$     & Depreciation of hard capital & $0.0206$  & BEA\\
            $\delta^i$           & Depreciation of ICT capital &  $0.0398$ & BEA \\
            $a$           & Hard capital share & $0.3$  & Standard value\ \\
            $b$           & ICT capital share &  $0.031$ & Oulton (2012) \\
            $\Gamma^c$           & Steady state growth rate of hard capital &  $1.0034$ & See below   \\
            $\Gamma^i$           & Steady state growth rate of ICT capital &  $1.0160$ & See below\\
            $\chi$           & Frisch elasticity &  $1$ & Standard value \\ 
			\hline
		\end{tabular}
		%\caption{$\beta$: Match average yield of 3month T-bill 1989:Q1-2017:Q1. $delta^i: match the higher of two BEA annual depreciation items that both correspond to ICT }
		\label{table:estimated_parameters}
	\end{center}
\end{table}

Get $(\Gamma^c, \Gamma^i)$ by solving the following model-implied steady state system

\begin{equation}\label{equation:system_param}
\begin{cases}
 \frac{p_t}{p_{t-1}}  = \frac{\Gamma^c}{\Gamma^i} \\
 \frac{c_t}{c_{t-1}}  = (\Gamma^c)^{\frac{1-b-\gamma}{1-a-b-\gamma}}  (\Gamma^i)^{\frac{b+\gamma}{1-a-b-\gamma}}
\end{cases}
\end{equation}

\hyperlink{IRmatching}{\beamerreturnbutton{Return}}
